---
title: "CBS buurt 2015"
author: "Marieke"
date: "December 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE)
```

# Initialization
## Loading required libraries
For raster and spatialpolygon classes we need the raster and rgdal packages. The projection of the data is in the Dutch RD-coordinate system. 
```{r initialization}
library(raster)
library(rgdal)
.libPaths("/usr/people/dirksen/R-20160725/x86_64-redhat-linux-gnu-library/3.3/")
library(lubridate)
library(caret)
library(doParallel)
registerDoParallel(8)
pro=CRS("+init=epsg:28992")
```

## Writing the required data to rasters
We want to calculate the mean values for each polygon for a average year, warm year and cold year. Herefore we take:
* 2012 average year
* 2003 warm year
* 1996 cold year

```{r select year}
# average<-stack(list.files("/nobackup/users/dirksen/Temperature/Temperature/Data/HARMONIE",pattern="file2012",full.names = TRUE))
# warm<-stack(list.files("/nobackup/users/dirksen/Temperature/Temperature/Data/HARMONIE/",pattern="file2003",full.names = TRUE))
# cold<-stack(list.files("/nobackup/users/dirksen/Temperature/Temperature/Data/HARMONIE/",pattern="file1996",full.names = TRUE))
# 
# proj4string(average)<-pro
# proj4string(warm)<-pro
# proj4string(cold)<-pro
# 
# writeRaster(average,"/nobackup/users/dirksen/Temperature/Temperature/Data/Results/average2012.grd")
# writeRaster(warm,"/nobackup/users/dirksen/Temperature/Temperature/Data/Results/warm2003.grd")
# writeRaster(cold,"/nobackup/users/dirksen/Temperature/Temperature/Data/Results/cold1996.grd")
```

```{r,eval=FALSE}

years<-seq(1996,2014,by=1)


foreach(i=1:length(years),.packages = c("raster","rgdal","lubridate")) %dopar% {
year<-years[i]
ptrn<-paste0("file",year)
filename<-paste0("grid",year,".grd")

r.st <- stack(list.files("/nobackup/users/dirksen/Temperature/Temperature/Data/HARMONIE",pattern=ptrn,full.names = TRUE))

proj4string(r.st)<-pro

writeRaster(r.st,paste0("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/allyears/",filename),overwrite=TRUE)
}
```
## Loading rasters
In the yearday_mean.grd are the values averaged over the period 1995 up to 2013. 
```{r}
st<-stack("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/yearday_mean.grd")
proj4string(st)<-pro

average<-stack("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/average2012.grd")
warm<-stack("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/warm2003.grd")
cold<-stack("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/cold1996.grd")
```

## CBS
First we read the unprojected data and look at the shapefile. 
> Missing the .dbf extention!!!

```{r cbs data}
cbs2014<-readOGR(dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/CSB2014",layer = "buurt_2014")
cbs2014_pro <- spTransform(cbs2014,crs(st))

summary(cbs2014)
```

The file yearday_mean contains the mean value for each day of the year averaged over the period 1995 until 2013 (366 layers). Here we use the complete cbs file from 2015. 
```{r loadCBS}
# cbs2015<-readOGR(dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/CBS2015/",layer = "Buurt_CBS_2015_v1")
# 
# 
# cbs2015_pro<-spTransform(cbs2015,crs(st))
```

# Big data analysis
## Day mean for all years
```{r meanallyears}
st.ext<-extract(st,cbs2014_pro,
                 fun=mean,nl=366,
                 method='bilinear',
                 sp=TRUE) #using this method the values from the 4 nearest cells are interpolated
writeOGR(st.ext,dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer ="klimatologie_buurt1995tm2013_daymean", driver="ESRI Shapefile")

library(doParallel)
library(foreach)
registerDoParallel(cores = 6)
df.alldays <- foreach(f=iter(nlayers(st)),.combine="cbind",.packages=c("raster")) %dopar% {
st.ext<-extract(st,cbs2014_pro,
                fun=mean,nl=1,layer=f,
                method='bilinear',
                sp=FALSE) #using this method the values from the 4 nearest cells are interpolated
}
```

## Daily temperature for 2012
```{r average2012}
average.ext<-extract(average,cbs2014_pro,
                 fun=mean,nl=366,
                 method='bilinear',
                 sp=TRUE) #using this method the values from the 4 nearest cells are interpolated
writeOGR(average.ext,dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer ="average_buurt2012", driver="ESRI Shapefile")
```

## Daily temperature for 2003
```{r warm2003}
warm.ext<-extract(warm,cbs2014_pro,
                 fun=mean,nl=366,
                 method='bilinear',
                 sp=TRUE) #using this method the values from the 4 nearest cells are interpolated
writeOGR(warm.ext,dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer ="warm_buurt2003", driver="ESRI Shapefile")
```

## Daily temperature for 1996
```{r cold1996}
cold.ext<-extract(cold,cbs2014_pro,
                 fun=mean,nl=366,
                 method='bilinear',
                 sp=TRUE) #using this method the values from the 4 nearest cells are interpolated
writeOGR(cold.ext,dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer ="cold_buurt2003", driver="ESRI Shapefile")
```

# A small test case for your code

## Reading the HARMONIE stack
Lets take an example with a single layer: the mean temperature in Kelvin from 1995 until 2013. The projection is added to the stack and set equal to the projection of the CBS shapefile. 
```{r HARMONIE data}
st<-stack("/nobackup/users/dirksen/Temperature/Temperature/Data/Results/alldays_mean.grd")
proj4string(st)<-pro
names(st)<-"Mean_Temperature_HARM"
cbs2015_pro<-spTransform(cbs2015,crs(st))

```

## Extracting the values for each polygon
Now the data is ready to be extracted. We want the mean of each polygon. The method is set to bilinear, using the 4 nearest grids. The default only uses the value of the grid falling with in the polygon. As a return we want the layer added to the shapefile and use sp=true. If you want a dataframe use sp=false. A new shapefile is generated with the command writeOGR. 
```{r extracting values}
time<-system.time(
st.ext2<-extract(st,cbs2014_pro,fun=mean,method='bilinear',sp=TRUE)
)
print(time)

writeOGR(st.ext2,dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer ="klimatologie_buurt1995tm2013", driver="ESRI Shapefile")

out<-readOGR(dsn="/nobackup/users/dirksen/Temperature/Temperature/Data/st_extracted_CBS/",layer="klimatologie_buurt1995tm2013")
```

