---
title: "Stations Europe"
author: "Marieke Dirksen"
date: "August 29, 2017"
output: 
   html_document:
            pandoc_args:  [
      "--title-prefix", "Foo",
      "--id-prefix", "Bar",
      "--number-sections"
      ]
            theme: yeti
            highlight: tango
            fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading the data

```{r stationdata, echo=FALSE,message=FALSE}
library(raster)
library(data.table)
library(leaflet)
library(rgdal)
library(maptools)
#library(spatstat) #problems with deldir installation, .rdb corrupt...
library(ggplot2)
library(ggmap)
library(GeoInterpolation)

subset_europe<-fread("/nobackup/users/dirksen/data/radiation_europe/qq_synoptical_stations_Jan2016.txt",header=TRUE)

stations_europe<-fread("/nobackup/users/dirksen/data/radiation_europe/qq_synoptical_stations_2016.txt",header=FALSE)
names(stations_europe)<-names(subset_europe)

cfg <- config::get(file = "/nobackup/users/dirksen/Temperature/GeoInterpolation/config/config.yml")
stations_europe<-data.frame(stations_europe)
```

# Visualization of the points 

## leaflet

```{r leaflet}
leafletplot<-stations_europe[!duplicated(stations_europe$syn_id),]

# m<-leaflet(data=leafletplot) %>% addTiles() %>% addMarkers(data=stations_europe,lng=~lon,lat=~lat,label=~as.character(syn_id))
# m
```

## ggplot2

```{r ggplot2}

map<-get_map(location="Europe",zoom=4)
ggmap(map)+geom_point(aes(x=lon,y=lat),data=leafletplot)

```

# Spatial pattern of the stations

* [online example](http://www.bias-project.org.uk/ASDARcourse/unit5_slides.pdf)

```{r spatialdensity}


spdf<-leafletplot
coordinates(spdf)<-~lon+lat
sp<-as(spdf,"SpatialPoints")

sp_ppp<-as(sp,"ppp")
bb <- boundingbox(sp_ppp)
ch <- convexhull.xy(sp_ppp)
rr <- ripras(sp_ppp)

sp_rr<- ppp(sp_ppp$x,sp_ppp$y,window=rr)

plot(sp_rr)

qr<-quadratcount(sp_ppp)
plot(qr)



spatial_density<-function(sigma){
crds<-list("x"=seq(min(sp_rr$x),max(sp_rr$x),by=0.1),"y"=seq(min(sp_rr$y),max(sp_rr$y),by=0.1))
k25<-density(sp_rr,sigma=sigma,xy=crds)
sgdf<-as(k25,"SpatialGridDataFrame")
r<-raster(sgdf)
crs(r)<-cfg$WGS84
return(r)
}

r25<-spatial_density(0.25)
r50<-spatial_density(0.50)
r75<-spatial_density(0.75)
r100<-spatial_density(1.00)

st<-stack(r25,r50,r75,r100)
names(st)<-c("sigma25","sigma50","sigma75","sigma100")
spplot(st,main="Kernel density estimation for different bandwidths")

```

## Spatial point patterns

Two different simulations:

* Complete spatial randomness, using G-function
* Distance from point to nearest event, using F-function

The output can be interpreted as:

* If the actual line is within the bounds of the theoretical line then the pattern is random
* If the line is below the pattern is clustered
* If the line is above the pattern is organized 

```{r spatialrandomness}
ex<-expression(runifpoint(n=sp_rr$n,win=rr))
res<-envelope(sp_rr,Kest,nsim=99,simulate=ex,verbose = FALSE,saveall=TRUE)
res
plot(res,main="CSR simulation") #clustered pattern

fes<-envelope(sp_rr,Fest,nsim=99,simulate=ex,verbose = FALSE,saveall=TRUE)
fes
plot(fes) # clustered pattern
```

# Interpolation

```{r interpolation, echo=FALSE,message=FALSE}
stations_europe$syn_date<-as.Date(stations_europe$syn_date)
stations_europe<-data.table(stations_europe)

# calculate the mean value for the days in Januari for all stations using the data.table package
# jan.meanwgs<-stations_europe[,.(mean=mean(qq)),by=.(syn_id,lat,lon)]
# coordinates(jan.meanwgs)<-~lon+lat
# proj4string(jan.meanwgs)<-cfg$WGS84
# jan.mean<-spTransform(jan.meanwgs,cfg$pro)
# jan.mean<-as(jan.mean,"SpatialPointsDataFrame")

#NOTE: ONLY THE FIRST DAY OF THIS DATASET HAS INFORMATION!
exampleday.wgs<-stations_europe[which(stations_europe$syn_date==unique(stations_europe$syn_date)[30]),]
exampleday.wgs$qq<-exampleday.wgs$qq/1000
coordinates(exampleday.wgs)<-~lon+lat
proj4string(exampleday.wgs)<-cfg$WGS84
exampleday<-spTransform(exampleday.wgs,cfg$pro)
exampleday<-as(exampleday,"SpatialPointsDataFrame")

# grid for europe
r<-raster()
extent(r)<-extent(exampleday.wgs)
res(r)<-0.25
values(r)<-1
crs(r)<-cfg$WGS84
r<-projectRaster(r,crs=cfg$pro)
sdf<-as(r,"SpatialGridDataFrame")


#IDW
# europe_int1<-interpolation_idw(groundstations = exampleday,
#                   variable = "qq",
#                   distshore.grd =  sdf
# )
# 
# r_int1<-raster(europe_int1$spatial)
# r_int1<-projectRaster(r_int1,crs=cfg$WGS84)
# pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(r_int1),
#   na.color = "transparent")
# 
# m1<-leaflet() %>% addTiles() %>% addRasterImage(r_int1,colors=pal,opacity=0.6) %>% addLegend(pal = pal, values = values(r_int1),title="Radiation [W/m2]") 
# 
# m1

# Some notes on the kriging interpolation:
# it doesn't work with WGS84 projection!
europe_int2<-interpolation_ordinarykriging(groundstations = exampleday,
                  variable = "qq",
                  grid_drift = sdf
)
```

```{r vis}
r_int2<-raster(europe_int2$spatial)
r_int2<-projectRaster(r_int2,crs=cfg$WGS84)
pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(r_int2),
  na.color = "transparent")

exampleday.df<-data.frame(exampleday.wgs)

pal2 <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), exampleday.df$qq,
  na.color = "transparent")

m2<-leaflet(data=exampleday.df) %>% addTiles() %>% addRasterImage(r_int2,colors=pal,opacity=0.6) %>% addLegend(pal = pal, values = values(r_int2),title="Radiation [KJ/m2]")  #%>% addCircleMarkers(lng=~lon,lat=~lat,label=~as.character(qq),fillColor=pal2,color=pal2)

m2
```