---
title: "Temperature Interpolation"
author: "Marieke Dirksen"
date: "Aug, 2017"
output: 
   html_document:
            pandoc_args:  [
      "--title-prefix", "Foo",
      "--id-prefix", "Bar",
      "--number-sections"
      ]
            theme: yeti
            highlight: tango
            fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = 'hide')
```

# Loading data

Important variables and paths are stored in the config.yml file. The station data used for the interpolation are documented data-sets which are loaded into memory. 

```{r,cache=TRUE,message=FALSE,echo=FALSE}
require(raster)
require(sp)
library(rgdal)
require(data.table)
require(config)
require(GeoInterpolation)
library(caret)
library(gstat)
library(leaflet)
library(htmltools)
library(rmarkdown)
library(rasterVis)

cfg <- config::get(file = "/nobackup/users/dirksen/Temperature/GeoInterpolation/config/config.yml")

#importing the datasets for the interpolation
devtools::load_all()
data("temperature_stationdata")
data("solar_radiation")

#check documentation 
?temperature_stationdata
?solar_radiation


```

## Subset data stations
```{r KNMIstations}
inputdata<-temperature_stationdata[which(temperature_stationdata$Datum==cfg$exampledatum),]




coordinates(inputdata)<-~RDN_X+RDN_Y
projection(inputdata)<-cfg$pro
```

```{r leaflet, , include=TRUE}
# For the location a fancy leaflet plot can be made
inputleaf<-spTransform(inputdata,cfg$WGS84)
inputleaf<-data.frame(inputleaf)

m<-leaflet(data=inputleaf) %>% addTiles() %>% addMarkers(lng=~RDN_X,lat=~RDN_Y,label=~htmlEscape(Locatie),labelOptions = labelOptions(noHide = T))

m

```

## Subset data grids

```{r HARMONIEdata}

examplefile<-paste0("file",cfg$exampledatum,"T123000Z.grd")
r.st <- stack(list.files(paste0(cfg$datapath,cfg$HARMONIEdata),pattern=examplefile,full.names = TRUE))
proj4string(r.st)<-cfg$pro

plot(r.st)

maskbuffer<-raster(paste0(cfg$datapath,cfg$nlbuffer))
proj4string(maskbuffer)<-cfg$pro
maskbuffer<-as(maskbuffer,"SpatialGridDataFrame")
plot(maskbuffer)

svf1km<-stack(list.files(path=cfg$svf1km,pattern = ".grd",full.names = TRUE))
names(svf1km)<-c("quantile 0.05","quantile 0.33","quantile 0.50","quantile 0.67","quantile 0.95")

svf100m<-stack(list.files(path=cfg$svf100m,pattern = ".grd",full.names = TRUE))
names(svf100m)<-c("quantile 0.05","quantile 0.15","quantile 0.33","quantile 0.50","quantile 0.67","quantile 0.95")



theme_set(theme_bw())
p<- gplot(svf1km) + geom_tile(aes(fill = value)) +
          facet_wrap(~ variable) +
          scale_fill_gradientn(colours = terrain.colors(100)) +
          #scale_fill_gradient(low = 'red', high = 'green') +
          coord_equal()
p

# ggsave(p,file="/usr/people/dirksen/Documents/Fotos/skyviewfactor/svf1km_quantiles.png")

p<- gplot(svf100m[[2]]) + geom_tile(aes(fill = value)) +
          facet_wrap(~ variable) +
          scale_fill_gradientn(colours = terrain.colors(100)) +
          #scale_fill_gradient(low = 'red', high = 'green') +
          coord_equal()
p

# ggsave(p,file="/usr/people/dirksen/Documents/Fotos/skyviewfactor/svf100m_quantiles.png")

distshore<-raster("/nobackup/users/dirksen/data/Temperature/KNMIstations/wn_distshore_001.asc")
proj4string(distshore)<-cfg$pro

distshore<-resample(distshore,r.st,method="ngb")
r.combined<-stack(r.st,distshore)
proj4string(r.combined)<-cfg$pro

r.combined<-as(r.combined,"SpatialGridDataFrame")
r.st <- as(r.st,"SpatialGridDataFrame")
distshore<-as(distshore,"SpatialGridDataFrame")

```

# Principle components
```{r pca}
pca_harm<-readRDS(cfg$pca_harmonie)
pca_siccs<-readRDS(cfg$pca_siccs)

proj4string(pca_harm$map)<-cfg$pro
proj4string(pca_siccs$map)<-cfg$pro

spdf_pca_harm<-as(pca_harm$map,"SpatialGridDataFrame")
spdf_pca_siccs<-as(pca_siccs$map,"SpatialGridDataFrame")
```

# Temperature interpolation 

The following interpolation methods are included: 

* inverse distance interpolation (IDW)
* thin plate splines (TPS)
* ordinary kriging (OK)
* kriging with external drift (KED)
* conditional gaussian simulations (option in KED, with standard nsim=4)

Input for the interpolation methods:

* groundstation data in spatial format
* variable you want to interpolate
* grid used either as drift or as mask, depending on the method
* maximum distance, used for different interpolation methods
* blocksize, also used for different interpolation methods
* overlay method, how you want to compare the points with the grid (either over or extract function)
* formula, used for kriging with drift
* conditional simulation, optional, used for kriging with drift

Output of the interpolations includes a list with the following objects:

* spatial
* cv
* stations
* statistical_summary_cv
* optional: conditional gaussian simulations

```{r temperature interpolation}

out.idw<-interpolation_idw(groundstations = inputdata, 
                           variable="Tn",
                           distshore.grd = maskbuffer)
out.tps<-interpolation_tps(groundstations = inputdata, 
                           variable="Tn" ,
                           distshore.grd = maskbuffer)
out.ok<-interpolation_ordinarykriging(groundstations = inputdata,
                                      variable="Tn",
                                      grid_drift=maskbuffer)
plot(out.ok$spatial)
spplot(out.idw$spatial,1)
spplot(out.tps$spatial,2)
#
```

# Comparisson between PCA HARMONIE and 1 day as trend
```{r pca vs daygrid}
f<-formula(paste("Tint ~",names(spdf_pca_harm)))

out.ked_pca<-interpolation_krigingdrift(groundstations = inputdata,
                                    formula=f,
                                    variable="Tn",
                                    grid_drift=spdf_pca_harm,
                                    method.overlay = 'extract',
                                    conditional.sim=FALSE)

f<-formula(paste("Tint ~",names(r.st)))

out.ked_1day<-interpolation_krigingdrift(groundstations = inputdata,
                                    formula=f,
                                    variable="Tn",
                                    grid_drift=r.st,
                                    method.overlay = 'extract',
                                    conditional.sim=FALSE)



plot(out.ked_pca$spatial)
plot(out.ked_1day$spatial)

out.ked_pca$statistical_summary_cv
out.ked_1day$statistical_summary_cv
```



# Other models

```{r radiation interpolation}
f<-formula("Tint ~ PC1 + PC2 + PC3 + PC4 + PC5")

lin.mod<-linear_model_caret(groundstations = inputdata,
                                    formula=f,
                                    variable="Tn",
                                    grid_prediction=spdf_pca_harm,
                                    method.overlay = 'extract'
                                    )


```

# Data preprocessing

In some cases the data can be skewed, or you use multiple data-sets with different oders of magnitude...You might wan't to preprocesses you data. Below the raster data is preprocessed, predicted and replaced in the raster object. 

```{r preprocess raster data}
# we can also interpolate the radiation data using the distance to sea grid and HARMONIE grid as trend...

#The raster data is preprocessed with the caret package, in this case using center, scale and boxcox functions
r.test<-r.combined
test<-preProcess(r.test@data,method=c("center","scale","BoxCox"))
newdata<-predict(test,r.test@data)

r.test@data<-newdata #replacing the data in the test raster with newdata

f<-formula(paste("Tint ~",paste(names(r.test),collapse = " + ")))

out.ked<-interpolation_krigingdrift(groundstations = radiation_data, 
                           variable="Q",
                           formula=f,
                           grid_drift=r.test,
                           method.overlay = 'extract')
plot(out.ked$spatial)
```




