---
title: "Ranger Interpolation"
author: "Marieke Dirksen"
date: "February 21, 2018"
output: 
   html_document:
            pandoc_args:  [
      "--title-prefix", "Foo",
      "--id-prefix", "Bar",
      "--number-sections"
      ]
            theme: yeti
            highlight: tango
            fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars,echo=FALSE,include=FALSE,cache=TRUE}
library(raster)
library(sp)
library(rgdal)
library(randomForest)
library(data.table)
library(config)
library(GeoInterpolation)
library(dplyr) # join function
library(caret)
library(lubridate)
library(GSIF)
library(ranger)
library(geoR)
library(spatialEco)
library(xgboost)

cfg <- config::get(file = "/nobackup/users/dirksen/Temperature/GeoInterpolation/config/config.yml")

#importing the datasets for the interpolation
devtools::load_all()
data("temperature_climate")
data("coords_aws")

temperature_climate<-temperature_climate[complete.cases(temperature_climate),]
fname<-"/nobackup/users/dirksen/data/Temperature/KNMIstations/temperature_auxiliary_over.txt"

data.aux.df<-fread(fname)
data.aux.df$yday<-yday(data.aux.df$Datum)
data.aux.df$originday<-as.numeric(as.POSIXct(data.aux.df$Datum))/(3600*24)
data.aux.df<-data.frame(data.aux.df)
data.aux.df<-subset(data.aux.df,select=c("STN","Datum","Tg","Height",
"Distsea","Population","Albedo","Roughness",
"Precipitation_monthly","NDVI","Radiation","yday","originday","RDN_X","RDN_Y"))

fm<-as.formula("Tg ~ Height + Distsea + 
               Population + Albedo + Roughness + 
               Precipitation_monthly + NDVI + Radiation +
               yday + originday")
#fm<-as.formula("Tg ~ Height + Distsea + Population + Albedo + Radiation")
data.aux.df<-data.aux.df[complete.cases(data.aux.df[,all.vars(fm)]),]


data.aux.sp<-data.aux.df
coordinates(data.aux.sp)<-~RDN_X+RDN_Y
projection(data.aux.sp)<-cfg$pro

domain<-read.asciigrid(paste0(cfg$datapath,cfg$nldistshore))
crs(domain)<-cfg$pro
domain.sp<-as(domain,"SpatialPixelsDataFrame")
```

# Ranger

## Spatial distances
Example of the spatial distances to all points. 

```{r sp.dist, cache=TRUE}
summary(data.aux.df)
data.aux.buf<-data.aux.df[unique(data.aux.df$STN),]
coordinates(data.aux.buf)<-~RDN_X+RDN_Y
projection(data.aux.buf)<-cfg$pro
grid.dist0<-GSIF::buffer.dist(data.aux.buf["Tg"],predictionDomain = domain.sp,as.factor(1:nrow(data.aux.buf)))
# r<-stack(grid.dist0)
# writeRaster(r,filename = "/nobackup/users/dirksen/data/auxcillary_NED/grid_buffer/grid_buffer.grd")
spplot(grid.dist0)
```

## Model

The model and predictions are based on this [link](https://github.com/thengl/GeoMLA#weighted-rfsp). The total uncertainty error of temperature measurements is estimated at 0.13. With the error the 'case.weights' are estimated. The yearly and long term trend, along with the albedo are found the most important variables for the model. 


```{r tuning, include=FALSE,cache=TRUE}
ov.lst<-list(data.aux.sp@data,over(data.aux.sp,grid.dist0))
data.aux.df<-do.call(cbind,ov.lst)
nms<-paste(names(data.aux.df)[14:length(names(data.aux.df))],collapse="+")

fm.buffer<-as.formula(paste("Tg ~ Height + Distsea + Population + Albedo + Roughness + Precipitation_monthly + NDVI + Radiation + yday + originday +",nms))

data.aux.df.sub<-data.aux.df[sample.int(size=5000,nrow(data.aux.df)),]

library(tuneRanger)
library(mlr)
Tg.task<-makeRegrTask(data=data.aux.df.sub[,-1:-2],target="Tg") #get the best mtry value
res = tuneRanger(Tg.task, num.trees = 1000, 
                  num.threads = 6, iters = 70,parameters = list(replace = FALSE,
                  respect.unordered.factors = TRUE)) # Problems with oder of the list, set to TRUE
```

```{r model,cache=TRUE}
print(res)
#tuneRF check https://www.rdocumentation.org/packages/randomForest/versions/4.6-12/topics/tuneRF 
m.Tg<-ranger(fm.buffer,data.aux.df, 
             num.trees=200, mtry=46,seed=1,
             quantreg = TRUE, importance='impurity')
#,case.weights=rep(1/(0.13),length(data.aux.df$STN))) #total uncertainty error of temperature measurements=0.13
print(m.Tg)
print(paste0("RMSE = ", round(sqrt(m.Tg$prediction.error),2)))

varimp<-data.table(m.Tg$variable.importance)
varimp$name<-names(m.Tg$variable.importance)
varimp<-setorder(varimp,-V1)
head(varimp,n=10)
# varimp$nr<-seq(1,length(varimp$name),1)
# ggplot(varimp,aes(V1,nr))+
#   geom_point() + 
#   geom_text(aes(label=name),hjust=-0.2,vjust=1) +
#   xlab("variable importance") +
#   ylab("ranking")
```

## Prediction

The code below loads the data for a random day and predicts the temperature. 

```{r prediction,cache=TRUE,include=FALSE,eval=TRUE}
datum<-as.Date("2006-11-01")
spdf_aux<-load_auxiliarydata(datum)
names(spdf_aux)<-c("Height","Distsea",
                   "Population","Albedo",
                   "Roughness","Precipitation_monthly",
                   "NDVI","Radiation","yday",
                   "originday",names(data.aux.df)[14:length(names(data.aux.df))])
p.Tg<-predict(m.Tg,spdf_aux@data,type="quantiles",quantiles=c(0.159,0.5,0.841)) #prediction

spdf_aux@data<-do.call(cbind, list(spdf_aux@data,
                                   data.frame(p.Tg$predictions[,2]),
                                   data.frame((p.Tg$predictions[,3]-p.Tg$predictions[,1])/2))) # add the data
gridded(spdf_aux)<-TRUE                   #spatialpixeldataframe
st<-stack(spdf_aux)                       #convert to raster

points<-data.aux.df[which(data.aux.df$Datum==datum),]
r.median<-st[[length(names(spdf_aux))-1]]
```

```{r plotting}
summary(points$Tg)
summary(r.median)
spplot(r.median,col.regions=terrain.colors(n=400))
```

```{r otherdate,cache=TRUE,include=FALSE,eval=TRUE}
datum<-as.Date("2010-12-20")
spdf_aux<-load_auxiliarydata(datum)
names(spdf_aux)<-c("Height","Distsea",
                   "Population","Albedo",
                   "Roughness","Precipitation_monthly",
                   "NDVI","Radiation","yday",
                   "originday",names(data.aux.df)[14:length(names(data.aux.df))])

p.Tg<-predict(m.Tg,spdf_aux@data,type="quantiles",quantiles=c(0.159,0.5,0.841)) #prediction

spdf_aux@data<-do.call(cbind, list(spdf_aux@data,
                                   data.frame(p.Tg$predictions[,2]),
                                   data.frame((p.Tg$predictions[,3]-p.Tg$predictions[,1])/2))) # add the data
gridded(spdf_aux)<-TRUE                   #spatialpixeldataframe
st<-stack(spdf_aux)                       #convert to raster

points<-data.aux.df[which(data.aux.df$Datum==datum),]
r.median<-st[[length(names(spdf_aux))-1]]

```


```{r plotting2}
summary(points$Tg)
summary(r.median)
spplot(r.median,col.regions=terrain.colors(n=400))
```

# xgboost

A test on a small subset...Working code but still poor results...

## Model
```{r xgmodel}
fitControl <- caret::trainControl(method = "repeatedcv", number = 10, repeats = 2, search = "random")
data.aux.df.sub<-data.aux.df[sample.int(size=500,nrow(data.aux.df)),]

I<-caret::createDataPartition(data.aux.df.sub$Tg,p=0.7,list = FALSE,times = 1)
train = data.aux.df.sub[I,]
test = data.aux.df.sub[-I,]
watchlist <- list(eval = test, train = train)

model <- caret::train(fm, data = data.aux.df.sub, method = "xgbTree", trControl = fitControl)
print(model)
```